---
title: "Stat 849 HW5"
author: "Lingyun Xiao"
date: '2022-11-30'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
solution <- read.csv("~/Library/CloudStorage/OneDrive-UW-Madison/Stat 849/solution.txt", sep="")
patient <- read.csv("~/Library/CloudStorage/OneDrive-UW-Madison/Stat 849/patient.txt", sep="")
```

## Stat 849 HW 5

### Q1

#### (a)

Population: The concentration of solution over time.

Sample: The 15 identical solutions.

A cause-and-effect relationship can naturally be established because we expect the concentration of the solution to decrease over time.

#### (b)

```{r}
lmod1 <- lm(conc ~ hour, data = solution)
summary(lmod1)
```

```{r}
plot(x = solution$hour, y = solution$conc,
     main = "Concentration vs Time", xlab = "Time", ylab = "Concentration")
abline(lmod1)
```

We can specify the simple linear regression model as follows: $\hat{Y} = \hat{\beta_0} + \hat{\beta_1} X$.

#### (c)

```{r}
lmod_null = lm(conc ~ 1, data = solution)
anova(lmod_null, lmod1)
```

We have $H_0: \beta_1 = 0$ and $H_1: \beta_1 \neq 0$. Based on the one-way ANOVA F-test, we reject $H_0$ and conclude that $\beta_1 \neq 0$.

We must make the following assumptions: Normality/Constant variance/Independence of errors; Correct model structure (linear); No unusual observations (outliers, leverages, influential points).

#### (d)

```{r}
plot(lmod1, which = c(1,2,4,5), cook.levels = 1)
```

We can observe the following abnormalities:

(a) Quadratic residuals vs fitted values. It violates that the assumption of a linear model.

(b) Short-tailed standardized residuals on normal Q-Q plot. It suggests that the errors are not normally distributed.

(c) Observation 2, 13, 15 have very high Cook's Distance, which suggests they are influential points.

Potential remedies: Transform the predictors and response.

#### (e)

I would suggest a Box-Cox transformation. First, all the observations are strictly positive. Also, I observe a nonlinear relationship between the predictors and the response.

#### (f)

```{r}
require(MASS)
boxcox(lmod1, lambda=seq(-0.25,0.75,by=0.05),plotit=T)
lambda_opt <- 0.04
trans_conc <- (solution$conc^(lambda_opt)-1)/(lambda_opt)
lmod2 <- lm(trans_conc ~ solution$hour)
```

```{r}
summary(lmod2)
```

```{r}
plot(x = solution$hour, y = trans_conc,
     main = "Trans_Concentration vs Time", xlab = "Time", ylab = "Trans_Concentration")
abline(lmod2)
```

#### (g)

It is clear that the $R^2$ value has significant improvement, and our transformed model satisfies the linearity assumption of the underlying model. Also, the residual errors have significant decrease.

### Q2

#### (a)

```{r}
lmod3 <- lm(y ~ x2, data = patient)
confint(lmod3)
```

We run a $t$-test to check whether the single coefficient $\beta_2 = 0$. We can see that the confidence interval lies below 0, so we reject $H_0: \beta_2 = 0$.

#### (b)

```{r}
lmod3_full <- lm(y ~ ., data = patient)
x = model.matrix(lmod3_full)[,-1]
library(faraway)
vif(x)
```

We can see that the VIF for all three coefficients are relatively low, so there is little evidence that the model suffers from multicollinearity issue.

#### (c)

```{r}
lmod4 <- lm(y ~ x1 + x3, data = patient)
anova(lmod4, lmod3_full)
```

We remove $X_2$ in the multiple linear regression and do a one-way ANOVA F-test. We conclude that we cannot reject $H_0: \beta_2 = 0$.

#### Q4

First, we generate data according to the true generative model.

```{r}
set.seed(123)
N = 50
x1 = rnorm(N, 0, 1)
x2 = rnorm(N, 0, 1)
x = cbind(1, x1, x2)
x = data.frame(x)
beta = matrix(c(6.33, -0.024, 0.15), ncol = 3) 
epsilon = rnorm(N, 0, 0.34^2)
Y = beta %*% t(x) + epsilon
testdata = x
testdata$Y = t(Y)

lmod5 = lm(Y ~ x1 + x2, data = testdata)
summary(lmod5)
```

Then, we introduce an outlier.

```{r}
p1 <- c(1, 0.5, 0.5, 20)
lmod6 <- lm(Y ~ x1 + x2, rbind(testdata, p1))
summary(lmod6)
```

```{r}
plot(lmod6, which = c(1,2,4,5), cook.levels = 1)
```

It is evident that the newly introduced outlier, point 51, significantly changes the fit. The coefficient $\beta_1$ becomes positive, and the t-value of both coefficients become much lower. Also, the $R^2$ value dramatically decreases to almost 0. From the plots, we can clearly see that point 51 has very high residuals vs fitted values, and it lies well above the normal Q-Q line. Point 51 also has a Cook's distance of nearly 0.5, suggesting that the outlier is also an influential point.
